---
title: "Artificial Intelligence Opinion Survey"
subtitle: "DATA 490 Independent Study"
author: "Mentor: Dr. Nicholas Dietrich\nAadarsha Gopala Reddy, Darren Lo, Ethan Love, Jose Mancilla, Santo Sumo"
date: "April 20, 2023"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: darkly
  urlcolor: blue
---

# 1. Load Data

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)

# Load data
edu <- read.csv("prolific_edu.csv")
health <- read.csv("prolific_health.csv")
retail <- read.csv("prolific_retail.csv")
tech <- read.csv("prolific_tech.csv")
```

# 2. Data Cleaning

```{r, message=FALSE, warning=FALSE}
# Combine data into one data frame after mutating Age to be one data type
edu <- edu %>% mutate(Age = as.character(Age))
health <- health %>% mutate(Age = as.character(Age))
retail <- retail %>% mutate(Age = as.character(Age))
tech <- tech %>% mutate(Age = as.character(Age))

combined <- bind_rows(edu, health, retail, tech)

# Keep only rows with United States as the Country of residence
combined <- combined %>% filter("Country of residence" == "United States")
```

# 3. Data Exploration

The columns in the dataset are:

- Submission id: Unique identifier for each submission

- Participant id: Unique identifier for each participant

- Status: Status of the submission

- Started at: Time the survey was started

- Completed at: Time the survey was completed

- Reviewed at: Time the survey was reviewed

- Archived at: Time the survey was archived

- Time taken: Time taken to complete the survey

- Completion code: Completion code for the survey

- Total approvals: Total approvals for the survey

- Employment-sector: Employment sector of the participant

- Age: Age of the participant

- Sex: Sex of the participant
- Ethnicity simplified: Ethnicity of the participant

- Country of birth: Birth country of the participant

- Country of residence: Residence country of the participant

- Nationality: Nationality of the participant

- Language: Language of the participant

- Student status: Is participant a student?

- Employment status: Is participant a student?
