---
title: "**Artificial Intelligence Opinion Survey**"
subtitle: "DATA 490 Independent Study"
author: "*Mentor:* Dr. Nicholas Dietrich\n\n *Researchers:* Aadarsha Gopala Reddy, Darren Lo,\n\nEthan Love, Jose Mancilla, Santo Sumo"
date: "April 20, 2023"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: darkly
  urlcolor: blue
---

# 1. Load Data

```{r load-data, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)

# Load data. Top row is column name.
edu <- read.csv("prolific_edu.csv")
health <- read.csv("prolific_health.csv")
retail <- read.csv("prolific_retail.csv")
tech <- read.csv("prolific_tech.csv")
qualtrics <- read.csv("qualtrics_data.csv")
```

# 2. Data Cleaning

```{r combine-prolific-data}
# Combine data into one data frame after mutating Age to be one data type
edu <- edu %>% mutate(Age = as.character(Age))
health <- health %>% mutate(Age = as.character(Age))
retail <- retail %>% mutate(Age = as.character(Age))
tech <- tech %>% mutate(Age = as.character(Age))

combined <- bind_rows(edu, health, retail, tech)
# export combined data to csv
# write.csv(combined, "combined_non_qualtrics.csv")
```

```{r add-filter-qualtrics-data}
# combine qualtrics and combined data using qualtrics data's
#     ProlificID column and combined data's Participant id
combined <- left_join(qualtrics, combined,
  by = c("ProlificID" = "Participant.id")
)
# rename "Duration..in.seconds." column to "Duration"
colnames(combined)[
  colnames(combined) == "Duration..in.seconds."
] <- "Duration"
# rename "Ethnicity.simplified" column to "Ethnicity"
colnames(combined)[
  colnames(combined) == "Ethnicity.simplified"
] <- "Ethnicity"
# remove Age.x and keep only Age.y column and rename Age.y to Age
combined <- combined %>%
  select(-Age.x) %>%
  rename(Age = Age.y)
# remove Status.x and Status.y columns
combined <- combined %>%
  select(-Status.x) %>%
  select(-Status.y)
# remove Finished, Progress, UserLanguage, DistributionChannel,
# Nationality, and Consent columns
combined <- combined %>%
  select(-Finished) %>%
  select(-Progress) %>%
  select(-UserLanguage) %>%
  select(-DistributionChannel) %>%
  select(-Nationality) %>%
  select(-Consent)


# remove rows where Submission.id is NA
combined <- combined %>%
  filter(!is.na(Submission.id))
# Keep only rows which say "United States" in "Country.of.residence" column
combined <- combined %>%
  filter(combined$Country.of.residence == "United States")

# Replace all cells that say "Information Technology" and
# "Science, Technology, Engineering & Mathematics" to "STEM/IT"
# in Employment.sector column
combined$Employment.sector[
  combined$Employment.sector == "Information Technology"
] <- "STEM/IT"
combined$Employment.sector[
  combined$Employment.sector == "Science, Technology, Engineering & Mathematics"
] <- "STEM/IT"

# Replace all cells in column EnhanceHurt that say
# "AI will neither enhance nor detract from my work" to "neither",
# "AI will enhance my work" to "enhance", and
# "AI will detract from my work" to "detract"
combined$EnhanceHurt[
  combined$EnhanceHurt == "AI will neither enhance nor detract from my work"
] <- "neither"
combined$EnhanceHurt[
  combined$EnhanceHurt == "AI will enhance my work"
] <- "enhance"
combined$EnhanceHurt[
  combined$EnhanceHurt == "AI will detract from my work"
] <- "detract"

# Replace all cells in column TimeEnergy that say
# "Save a lot of time" to "lot of time",
# "Save little time" to "little time",
# "Save no time" to "no time",
# "Save some time" to "some time", and
# "Neutral" to "neutral"
combined$TimeEnergy[
  combined$TimeEnergy == "Save a lot of time"
] <- "lot of time"
combined$TimeEnergy[
  combined$TimeEnergy == "Save little time"
] <- "little time"
combined$TimeEnergy[
  combined$TimeEnergy == "Save no time"
] <- "no time"
combined$TimeEnergy[
  combined$TimeEnergy == "Save some time"
] <- "some time"
combined$TimeEnergy[
  combined$TimeEnergy == "Neutral"
] <- "neutral"

# export data to csv
# write.csv(combined, "combined_qualtrics.csv")
```

```{r keep-attentive-responses}
# Keep only rows which say "Compose an email" in "Attention" column
combined <- combined %>% filter(combined$Attention == "Compose an email")
# remove Attention column
combined <- combined %>% select(-Attention)
# export data to csv
# write.csv(combined, "combined_qualtrics_attentive.csv")
```
\newpage

# 3. Data Exploration

The columns in the dataset are:

- *StartDate* - Date and time survey was started

- *EndDate* - Date and time survey was completed

- *IPAddress* - IP address of participant

- *Duration* - Duration of survey in seconds

- *RecordedDate* - Date and time survey was recorded

- *ResponseId* - Response ID

- *LocationLatitude* - Participant's location latitude

- *LocationLongitude* - Participant's location longitude

- *ProlificID* - Identification of the response on Prolific

- *Gender* - Gender of the participant

- *Education* - Education level of the participant

- *Salary* - Salary of the participant

- *AIKnowledge* - Knowledge of AI of the participant

- *UsedAI* - Whether the participant has used AI

- *TimeEnergy* - How much time and energy AI has saved the participant

- *SimilarTasks* - How much of the participant's tasks they think AI can do

- *EnhanceHurt* - Whether the participant thinks AI can enhance or hurt their work efficiency.

- *Comments* - Comments from the participant

- *Submission.id* - Submission ID

- *Started.at* - Date and time survey was started

- *Completed.at* - Date and time survey was completed

- *Reviewed.at* - Date and time survey was reviewed

- *Archived.at* - Date and time survey was archived

- *Time.taken* - Duration of survey in seconds

- *Completion.code* - Completion code

- *Total.approvals* - Total number of approvals

- *Employment.sector* - Employment sector

- *Age* - Age of the participant

- *Sex* - Sex of the participant

- *Ethnicity* - Ethnicity of the participant

- *Country.of.birth* - Country of birth of the participant

- *Country.of.residence* - Country of residence of the participant

- *Language* - Language of the participant

- *Student.status* - Whether the participant is a student

- *Employment.status* - Whether the participant is employed
\newpage

# 4. Data Analysis

## 4.1. Average amount of time to complete survey

```{r avg-time}
# Create a new data frame with only the columns we need
avg_time <- combined %>% select(Duration)
# Remove rows where Duration is NA
avg_time <- avg_time %>% filter(!is.na(Duration))
# Calculate the average time taken to complete the survey
avg_time <- avg_time %>% summarise(avg_time = mean(Duration))
# Convert to <x> minutes and <y> seconds
avg_time$avg_time <- avg_time$avg_time / 60
avg_time$avg_time <- paste0(
  floor(avg_time$avg_time),
  " minutes and ",
  round((avg_time$avg_time - floor(avg_time$avg_time)) * 60),
  " seconds"
)
# Print the average time taken to complete the survey
avg_time
```

The average time taken to complete the survey is 2 minutes and 13 seconds. 
\newpage

## 4.2. EnhanceHurt vs. Sector

```{r enhancehurt-vs-sector, message = FALSE, warning = FALSE}
# Create a new data frame with only the columns we need
enhancehurt_vs_sector <- combined %>%
  select(EnhanceHurt, Employment.sector)
```

```{r enhancehurt-vs-sector-edit, message = FALSE, warning = FALSE, echo = FALSE}
# Remove rows where Employment.sector is NA
enhancehurt_vs_sector <- enhancehurt_vs_sector %>%
  filter(!is.na(Employment.sector))
# Remove rows where EnhanceHurt is NA
enhancehurt_vs_sector <- enhancehurt_vs_sector %>%
  filter(!is.na(EnhanceHurt))
```

```{r enhancehurt-vs-sector-plot, message = FALSE, warning = FALSE}
# For each sector, calculate the proportion of
# respondents who think AI can enhance their work efficiency
enhancehurt_vs_sector <- enhancehurt_vs_sector %>%
  group_by(Employment.sector, EnhanceHurt) %>%
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
# Visualize using different histogram for each sector.
# Show the number of respondents inside each bar.
# make legend bottom. Wrap x axis labels without changing plot size.
enhancehurt_vs_sector_plot <- ggplot(
  enhancehurt_vs_sector,
  aes(
    x = Employment.sector,
    y = prop, fill = factor(EnhanceHurt,
      levels = c("detract", "neither", "enhance")
    )
  ),
  ylim = c(0, 1)
) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count),
    position = position_dodge(width = 1),
    vjust = 1.5
  ) +
  labs(
    x = "Employment Sector", y = "Proportion of Respondents",
    fill = "Enhance or Hurt"
  ) +
  ggtitle("Proportion of Respondents Who Think AI Can\n
  Enhance Their Work Efficiency by Employment Sector") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

enhancehurt_vs_sector_plot
# save to png
ggsave("enhancehurt_vs_sector_plot.png",
  width = 10, height = 6, units = "in", dpi = 300
)
```
\newpage

## 4.3. EnhanceHurt vs. Education

```{r enhancehurt-vs-education, message = FALSE, warning = FALSE}
# Create a new data frame with only the columns we need
enhancehurt_vs_education <- combined %>%
  select(EnhanceHurt, Education)
```

```{r enhancehurt-vs-education-edit, message = FALSE, warning = FALSE, echo = FALSE}
# Remove rows where Education is NA
enhancehurt_vs_education <- enhancehurt_vs_education %>%
  filter(!is.na(Education))
# Remove rows where EnhanceHurt is NA
enhancehurt_vs_education <- enhancehurt_vs_education %>%
  filter(!is.na(EnhanceHurt))
```

```{r enhancehurt-vs-education-plot, message = FALSE, warning = FALSE}
# For each education level, calculate the proportion of
# respondents who think AI can enhance their work efficiency
enhancehurt_vs_education <- enhancehurt_vs_education %>%
  group_by(Education, EnhanceHurt) %>%
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
# Visualize using different histogram for each education level.
# Show the number of respondents on each bar.
enhancehurt_vs_education_plot <- ggplot(
  enhancehurt_vs_education,
  aes(
    x = Education, y = prop, fill = factor(EnhanceHurt,
      levels = c("detract", "neither", "enhance")
    )
  ),
  ylim = c(0, 1)
) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count),
    position = position_dodge(width = 1),
    vjust = 1.5
  ) +
  labs(
    x = "Education Level", y = "Proportion of Respondents",
    fill = "Enhance or Hurt"
  ) +
  ggtitle("Proportion of Respondents Who Think AI Can\n
  Enhance Their Work Efficiency by Education Level") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "bottom")

enhancehurt_vs_education_plot
# save to png
ggsave("enhancehurt_vs_education_plot.png",
  width = 10, height = 6, units = "in", dpi = 300
)
```
\newpage

## 4.4. EnhanceHurt vs. Salary

```{r enhancehurt-vs-salary, message = FALSE, warning = FALSE}
# Create a new data frame with only the columns we need
enhancehurt_vs_salary <- combined %>% select(EnhanceHurt, Salary)
```

```{r enhancehurt-vs-salary-edit, message = FALSE, warning = FALSE, echo = FALSE}
# Remove rows where Salary is NA
enhancehurt_vs_salary <- enhancehurt_vs_salary %>%
  filter(!is.na(Salary))
# Remove rows where EnhanceHurt is NA
enhancehurt_vs_salary <- enhancehurt_vs_salary %>%
  filter(!is.na(EnhanceHurt))

# remove rows where Salary is empty
enhancehurt_vs_salary <- enhancehurt_vs_salary %>%
  filter(Salary != "")
```

```{r enhancehurt-vs-salary-plot, message = FALSE, warning = FALSE}
# Create a new dataframe with the proportion of respondents who
# think AI can enhance their work efficiency for each salary group
enhancehurt_vs_salary <- enhancehurt_vs_salary %>%
  group_by(Salary, EnhanceHurt) %>%
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
# Visualize using different histogram for each salary group.
# Show the number of respondents on each bar.
enhancehurt_vs_salary_plot <- ggplot(
  enhancehurt_vs_salary,
  aes(
    x = Salary, y = prop, fill = factor(EnhanceHurt,
      levels = c("detract", "neither", "enhance")
    )
  ),
  ylim = c(0, 1)
) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count),
    position = position_dodge(width = 1),
    vjust = 1.5
  ) +
  labs(
    x = "Salary Group",
    y = "Proportion of Respondents",
    fill = "Enhance or Hurt"
  ) +
  ggtitle("Proportion of Respondents Who Think AI Can\n
  Enhance Their Work Efficiency by Salary Group") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "bottom")

enhancehurt_vs_salary_plot
# save to png
ggsave("enhancehurt_vs_salary_plot.png",
  width = 10, height = 6, units = "in", dpi = 300
)
```
\newpage

## 4.5. EnhanceHurt vs. Gender

```{r enhancehurt-vs-gender, message = FALSE, warning = FALSE}
# Create a new data frame with only the columns we need
enhancehurt_vs_gender <- combined %>% select(EnhanceHurt, Gender)
```

```{r enhancehurt-vs-gender-edit, message = FALSE, warning = FALSE, echo = FALSE}
# Remove rows where Gender is NA
enhancehurt_vs_gender <- enhancehurt_vs_gender %>%
  filter(!is.na(Gender))
# Remove rows where EnhanceHurt is NA
enhancehurt_vs_gender <- enhancehurt_vs_gender %>%
  filter(!is.na(EnhanceHurt))

# remove rows where Gender is empty
enhancehurt_vs_gender <- enhancehurt_vs_gender %>%
  filter(Gender != "Prefer to self-describe:")
```

```{r enhancehurt-vs-gender-plot, message = FALSE, warning = FALSE}
# Create a new dataframe with the proportion of respondents who
# think AI can enhance their work efficiency for each gender
enhancehurt_vs_gender <- enhancehurt_vs_gender %>%
  group_by(Gender, EnhanceHurt) %>%
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
# Visualize the proportion of respondents who think AI can
# enhance or hurt, based on gender, using a bar chart.
# Show the number of respondents inside each bar.
# make legend bottom. Wrap x axis labels without changing plot size.
enhancehurt_vs_gender_plot <- ggplot(
  enhancehurt_vs_gender,
  aes(
    x = Gender,
    y = prop, fill = factor(EnhanceHurt,
      levels = c("detract", "neither", "enhance")
    )
  ),
  ylim = c(0, 1)
) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count),
    position = position_dodge(width = 1),
    vjust = 1.5
  ) +
  labs(
    x = "Gender", y = "Proportion of Respondents",
    fill = "Enhance or Hurt"
  ) +
  ggtitle("Proportion of Respondents Who Think AI Can\n
  Enhance Their Work Efficiency by Gender") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

enhancehurt_vs_gender_plot
# save to png
ggsave("enhancehurt_vs_gender_plot.png",
  width = 10, height = 6, units = "in", dpi = 300
)
```
\newpage

## 4.6. EnhanceHurt vs. Ethnicity

```{r enhancehurt-vs-country, message = FALSE, warning = FALSE}
# Create a new data frame with only the columns we need
enhancehurt_vs_ethnicity <- combined %>% select(EnhanceHurt, Ethnicity)
```

```{r enhancehurt-vs-country-edit, message = FALSE, warning = FALSE, echo = FALSE}
# Remove rows where Country is NA
enhancehurt_vs_ethnicity <- enhancehurt_vs_ethnicity %>%
  filter(!is.na(Ethnicity))
# Remove rows where EnhanceHurt is NA
enhancehurt_vs_ethnicity <- enhancehurt_vs_ethnicity %>%
  filter(!is.na(EnhanceHurt))

# Remove rows where Country is empty
enhancehurt_vs_ethnicity <- enhancehurt_vs_ethnicity %>%
  filter(Ethnicity != "")
# Remove the Ethnicity which says "DATA_EXPIRED"
enhancehurt_vs_ethnicity <- enhancehurt_vs_ethnicity %>%
  filter(enhancehurt_vs_ethnicity$Ethnicity != "DATA_EXPIRED")
```

```{r enhancehurt-vs-country-plot, message = FALSE, warning = FALSE}
# Create a new dataframe with the proportion of respondents who
# think AI can enhance their work efficiency for each country
enhancehurt_vs_ethnicity <- enhancehurt_vs_ethnicity %>%
  group_by(Ethnicity, EnhanceHurt) %>%
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
# Visualize the proportion of respondents who think AI can
# enhance or hurt, based on country, using a bar chart.
# Show the number of respondents inside each bar.
# make legend bottom. Wrap x axis labels without changing plot size.
enhancehurt_vs_ethnicity_plot <- ggplot(
  enhancehurt_vs_ethnicity,
  aes(
    x = Ethnicity,
    y = prop, fill = factor(EnhanceHurt,
      levels = c("detract", "neither", "enhance")
    )
  ),
  ylim = c(0, 1)
) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count),
    position = position_dodge(width = 1),
    vjust = 1.5
  ) +
  labs(
    x = "Ethnicity",
    y = "Proportion of Respondents",
    fill = "Enhance or Hurt"
  ) +
  ggtitle("Proportion of Respondents Who Think AI Can\n
  Enhance Their Work Efficiency by Ethnicity") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

enhancehurt_vs_ethnicity_plot

# save to png
ggsave("enhancehurt_vs_ethnicity_plot.png",
  width = 10, height = 6, units = "in", dpi = 300
)
```
\newpage

## 4.7. TimeEnergy vs. Education

```{r timeenergy-vs-education, message = FALSE, warning = FALSE}
# Create a new data frame with only the columns we need
timeenergy_vs_education <- combined %>% select(TimeEnergy, Education)
```

```{r timeenergy-vs-education-edit, message = FALSE, warning = FALSE, echo = FALSE}
# Remove rows where Education is NA
timeenergy_vs_education <- timeenergy_vs_education %>%
  filter(!is.na(Education))
# Remove rows where TimeEnergy is NA
timeenergy_vs_education <- timeenergy_vs_education %>%
  filter(!is.na(TimeEnergy))

# Remove rows where Education is empty
timeenergy_vs_education <- timeenergy_vs_education %>%
  filter(Education != "")
```

```{r timeenergy-vs-education-plot, message = FALSE, warning = FALSE}
# Create a new dataframe with the proportion of respondents who
# think AI can save their time and energy for each education level
timeenergy_vs_education <- timeenergy_vs_education %>%
  group_by(Education, TimeEnergy) %>%
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
# Visualize the proportion of respondents who think AI can
# save their time and energy, based on education level, using a bar chart.
# Show the number of respondents inside each bar.
# make legend bottom. Wrap x axis labels without changing plot size.
# order the bars as "no time", "little time", "some time", "lot of time", and "neutral".
timeenergy_vs_education_plot <- ggplot(
  timeenergy_vs_education,
  aes(
    x = Education,
    y = prop, fill = factor(TimeEnergy,
      levels = c("no time", "little time", "some time", "lot of time", "neutral")
    )
  ),
  ylim = c(0, 1)
) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count),
    position = position_dodge(width = 1),
    vjust = 1.5
  ) +
  labs(
    x = "Education", y = "Proportion of Respondents",
    fill = "Time and Energy"
  ) +
  ggtitle("Proportion of Respondents Who Think AI Can\n
  Save Their Time and Energy by Education Level") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

timeenergy_vs_education_plot

# save to png
ggsave("timeenergy_vs_education_plot.png",
  width = 10, height = 6, units = "in", dpi = 300
)
```
\newpage

## 4.8. SimilarTasks vs. Age

```{r similartasks-vs-age, message = FALSE, warning = FALSE}
# Create a new data frame with only the columns we need
similartasks_vs_age <- combined %>% select(SimilarTasks, Age)
```

```{r similartasks-vs-age-edit, message = FALSE, warning = FALSE, echo = FALSE}
# Remove the Ages which say "DATA_EXPIRED" and "923"
similartasks_vs_age <- similartasks_vs_age %>%
  filter(similartasks_vs_age$Age != "DATA_EXPIRED")
similartasks_vs_age <- similartasks_vs_age %>%
  filter(similartasks_vs_age$Age != "923")
# Remove rows where Age is NA
similartasks_vs_age <- similartasks_vs_age %>%
  filter(!is.na(Age))
# Remove rows where SimilarTasks is NA
similartasks_vs_age <- similartasks_vs_age %>%
  filter(!is.na(SimilarTasks))

# Change age rows which say "79" and "80" to "73"
similartasks_vs_age$Age[similartasks_vs_age$Age == "79"] <- "73"
similartasks_vs_age$Age[similartasks_vs_age$Age == "80"] <- "73"

# Convert Age to numeric
similartasks_vs_age$Age <- as.numeric(as.character(similartasks_vs_age$Age))
# group ages by 5 years
similartasks_vs_age$Age <- cut(similartasks_vs_age$Age,
  breaks = seq(18, 80, by = 10)
)
# remove NA
similartasks_vs_age <- similartasks_vs_age %>% filter(!is.na(Age))

# create a new dataframe with the proportion of respondents who think AI can enhance their work efficiency for each age group
similartasks_vs_age <- similartasks_vs_age %>%
  group_by(Age, SimilarTasks) %>%
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
```

```{r similartasks-vs-age-plot, message = FALSE, warning = FALSE}
# Visualize the proportion of respondents who think AI can
# do similar tasks, based on age group, using a bar chart.
# Show the number of respondents inside each bar.
# make legend bottom. Wrap x axis labels without changing plot size.
# order the bars as "Cannot perform any tasks", "Can perform only a few tasks",
# "Can perform many tasks", "Can perform all tasks".
# force the y axis to start at 0 and end at 1.
similartasks_vs_age_plot <- ggplot( 
  similartasks_vs_age,
  aes(
    x = Age,
    y = prop, fill = factor(SimilarTasks,
      levels = c("Cannot perform any tasks", "Can perform only a few tasks",
      "Can perform many tasks", "Can perform all tasks")
    )
  ),
  ylim = c(0, 1)
) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count),
    position = position_dodge(width = 1),
    vjust = 1.5
  ) +
  labs(
    x = "Age", y = "Proportion of Respondents",
    fill = "Similar Tasks"
  ) +
  ggtitle("Proportion of Respondents Who Think AI Can\n
  Do Similar Tasks by Age Group") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

similartasks_vs_age_plot

# save to png
ggsave("similartasks_vs_age_plot.png",
  width = 10, height = 6, units = "in", dpi = 300
)
```
\newpage